Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-11-26 2017-Xudong-SMT/checker.tex
Update checker.tex
Add the draft of description of the format translation between pNet and Z3.

\documentclass[a4paper]{llncs}

@@

\def\AlgT{\mathcal{T}}
\def\AlgE{\mathcal{E}}
\def\AlgA{\mathcal{A}}
\def\AlgAS{\mathcal{A}_S}
\def\AlgB{\mathcal{B}}
\def\AlgI{\mathcal{I}}
\newcommand{\set}[1]{\overline{#1}}
\newcommand{\Pred}{\symb{Pred}}
\newcommand{\Post}{\symb{Post}}
%\usepackage[math]{cellspace}
%\setlength\cellspacetoplimit{ 37pt}
%\setlength\cellspacebottomlimit{18pt}

@@

% Macros for the SOS rules and proof trees:
%\newcommand\openrule[2]{\redinfer{#1}{#2}}
\newcommand\openrule[2]{\inferrule*[myfraction=\reddottedrule,center]{#1}{#2}}
%\newcommand\openrule[2]{\inferrule*{#1}{#2}}
%\newcommand\ostate[1]{\triangleleft{\;#1\;}\triangleright}
\newcommand\ostate[1]{\triangleleft{#1}\triangleright}
\newcommand{\sm}[1]{\mbox{\boldmath\small #1}}
\DeclareMathOperator{\card}{card}
\DeclareMathOperator{\Flat}{Flat}
\renewcommand{\P}{\mathcal P}

\begin{document}
\section{Type system}
The SMT solver Z3 already has conservative type rules and checks type safety in Horn logic. When translating the pNets into SMT-LIB language, pNets should also make sure that
the term matching other term in SMT-LIB language will be correct in type checking. In next two section, we give out the presentation of the pNets and the type rules of them to ensure the type safety. 
\subsection{Algebra Presentation}
We have a set of basic algebra presentation for the pNet, including three basic sorts $Bool$, $Action$ and $Int$ and their own operators(constants as operators with no argument).
\begin{table}\caption{\leftline{Algebra Presentation}}
	\begin{tabular}{p{4cm}p{8cm}}
		\hline\specialrule{0em}{1pt}{1pt}
		Sort & Operator \\\specialrule{0em}{1pt}{1pt}
		\hdashline\specialrule{0em}{3pt}{3pt}
		Bool    			& $=,\ \ne,\ \land,\ \lor,\ \texttt{true},\ \texttt{false},\ \texttt{FUN}$ 								\\\specialrule{0em}{1pt}{1pt}
		Action 			& $=,\ \ne,\ \texttt{FUN}$ 																	\\\specialrule{0em}{1pt}{1pt}
		Int 				& $=,\ \ne,\ - \texttt{(unary)},\ +,\ - \texttt{(binary)},\ \times,\ \div,\ \texttt{Nat},\ \texttt{FUN}$ 			\\\specialrule{0em}{1pt}{1pt}
		\hline
	\end{tabular}
\end{table}

@@

>>>>

\documentclass[a4paper]{lncs/llncs}

@@

\def\AlgT{\mathcal{T}}
\def\AlgE{\mathcal{E}}
\def\AlgA{\mathcal{A}}
\def\AlgAS{\mathcal{A}_S}
\def\AlgB{\mathcal{B}}
\def\AlgI{\mathcal{I}}
%\newcommand{\set}[1]{\overline{#1}}
\newcommand{\Pred}{\symb{Pred}}
\newcommand{\Post}{\symb{Post}}
%\usepackage[math]{cellspace}
%\setlength\cellspacetoplimit{ 37pt}
%\setlength\cellspacebottomlimit{18pt}

@@

% Macros for the SOS rules and proof trees:
%\newcommand\openrule[2]{\redinfer{#1}{#2}}
\newcommand\openrule[2]{\inferrule*[myfraction=\reddottedrule,center]{#1}{#2}}
%\newcommand\openrule[2]{\inferrule*{#1}{#2}}
%\newcommand\ostate[1]{\triangleleft{\;#1\;}\triangleright}
\newcommand\ostate[1]{\triangleleft{#1}\triangleright}
\newcommand{\sm}[1]{\mbox{\boldmath\small #1}}
%\DeclareMathOperator{\card}{card}
%\DeclareMathOperator{\Flat}{Flat}
\renewcommand{\P}{\mathcal P}

\begin{document}

This document presumes the definition of Term algebras and
parameterized actions from the pNet
theory definition, that we will not repeat here.

\section{Translations from pNets to SMTLib}
In order to submit satisfiability problems to Z3 (for the predicates
in open transitions), we need to generate SMTlib programs, from the
pNet Algebra presentation and predicates.
More precisely, we need to translate to SMTlib:
\begin{itemize}
  \item for a given pNet: the presentation of the action algebra
    (sorts, operators, constants), 
    \item for each open transition: the declaration of variables, and
      the predicate (including action expressions).
\end{itemize}

During this translation, in order to guarantee that the generated code
will cause no runtime errors during parsing and execution, we need
to ensure that all objects used in the SMTlib code are properly
declared, and that they are correctly typed.

\section{Type system}
%% The SMT solver Z3 already has conservative type rules and checks type safety in Horn logic. When translating the pNets into SMT-LIB language, pNets should also make sure that
%% the term matching other term in SMT-LIB language will be correct in
%% type checking. In next two section, we give out the presentation of
%% the pNets and the type rules of them to ensure the type safety.
The type system in pNets is quite simple, the expressions are built
from variables, constants, and operators of a many-sorted algebra, and
we only need simple first-order polymorphic types.
Additionaly, for convenience, we want to introduce some ``generic''
operators in the algebra, with an overloading mechanism; as these do
not exist in SMTlib, we will have a specific translation mechanism.

Note that in principle, such an algebra correspond to a given
high-level language (e.g. a process algebra), and that the algebra
presentation will be defined once and for all in the framework of the
pNet semantics of each specific language.

\subsection{Algebra Presentations}

We have a minimal, predefined algebra presentation for all pNets, including
three basic sorts $Bool$, $Action$ and $Int$ and their
operators. Table \ref{Table:predefinedSorts} defines these elements.

\begin{table}\caption{\label{Table:predefinedSorts}\leftline{Algebra Presentation: predefined Sorts
  and Operators}}
	\begin{tabular}{p{4cm}p{8cm}}
		\hline\specialrule{0em}{1pt}{1pt}
		Sort & Operator
                \\\specialrule{0em}{1pt}{1pt}
		\hline\specialrule{0em}{3pt}{3pt}
		Bool    			&
                $\land,\ \lor,\ \texttt{true},\ \texttt{false}$
                \\\specialrule{0em}{1pt}{1pt} 
		Action 			& \textsl{no predefined operator}
                \\\specialrule{0em}{1pt}{1pt}
		Int 				& $\ - \texttt{(unary)},\ +,\ -
                \texttt{(binary)},\ \times,\ \div,\ {0, 1, ...} \in \texttt{Nat}$
                \\\specialrule{0em}{1pt}{1pt}
                \textsl{for any sort} & $=,\ \ne$
		\\\hline
	\end{tabular}
\end{table}

For a given language, or for a given use-case, the designer can
declare more sorts and operators, using our pNet API.
As an example, the CCS action algebra include:

@@

\paragraph{Translation} from pNets informations into SMT-LIB language format is corresponded with the mapping rules.
\subsubsection{Sort}
There are several basic Sorts already defined in the presentation and users can declare more customized Sorts. In addition, Bool and Int type is also Z3 default datatypes. Sorts only provide their names to the datatype constructor in Z3, the more content will be added according to the Operators in the presentation.\\\\
\centerline{$\texttt{(declare-datatypes\ ()\ (($SortName$\ ...\ )))}$}
%(declare-datatypes () ((SortName ... )))
\subsubsection{FUN function}
The signature of the FUN operators are defined by users. But they still have a fixed format.\\\\
\centerline{$\texttt{FUN} :: Type_1, ... ,Type_n \longrightarrow Type$}\\\\
%(declare-fun FUN (TypeName_1, ... ,TypeName_n) TypeName)
The types of all the arguments are given to the Z3, so does the return type, but written at another place.  We use a suffix after the "FUN" name to distinct different FUN operators.\\\\
\centerline{$\texttt{(declare-fun\ FUN\_Type1...\_TypeN\_Type\ ($Type_1$\ ... \ $Type_n$)\ $Type$)}$}
